---
title: "DATA 621 HW 4"
author: "IvanTikhonov, Seung min song, Alice Friedman, Anthony A"
date: "2023-03-08"
output: html_document
---

Overview
In this homework assignment, you will explore, analyze and model a data set containing approximately 8000
records representing a customer at an auto insurance company. Each record has two response variables. The
first response variable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero
means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero
if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero.

Data Exploration
```{r libraries, include=FALSE, warning=FALSE, class.output="scroll-less"}
knitr::opts_chunk$set(echo = TRUE)
library(e1071)
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(FactoMineR)
library(VIFCP)
library(knitr)
library(kableExtra)
library(Hmisc)
library(pROC)
library(binr)
library(mice)
```

```{r, include=FALSE, echo=FALSE}

rawData = read.csv("C:/Users/SeungminSong/Downloads/621R/insurance-evaluation-data.csv")
#rawTestData = read.csv("https://raw.githubusercontent.com/Anthogonyst/AnalyticsModeling/master/data/insurance_training_data.csv?token=GHSAT0AAAAAACAEYZJ5FGMCRFNVCGWAPOIKZBTAXSQ")

```


```{r}
head(rawData)
sapply(rawData, class)
```

## Tidy Data

After carefully examining the data in the rawData data frame, it was found that certain variables contained unnecessary characters such as **$**, **z_**, **<** which could potentially hinder further analysis. To address this issue, regular expression (regex) was used to remove these unwanted characters from the variables. By doing so, the data is now cleaner and more suitable for analysis. 
```{r}
# create a vector of column names in rawData
col_names <- names(rawData)

# loop through each column and remove '$' and 'z_' and '<' and ','
for (col in col_names) {
   rawData[[col]] <- gsub("\\$|z_|<|,", "", rawData[[col]])
}
head(rawData)
sapply(rawData, class)
```

change "no" to 0 and "yes" to 1 for the columns **MSTATUS**, **RED_CAR**, **REVOKED**, and **PARENT1**. 

Convert the following columns to integer data type: KIDSDRIV, AGE, HOMEKIDS, YOJ, INCOME, HOME_VAL, BLUEBOOK, TIF, OLDCLAIM, CLM_FREQ, MVR_PTS, and CAR_AGE.

```{r echo=FALSE}
rawData$MSTATUS <- as.numeric(rawData$MSTATUS == "Yes")
rawData$RED_CAR <- as.numeric(rawData$RED_CAR == "yes")
rawData$REVOKED <- as.numeric(rawData$REVOKED == "Yes")
rawData$PARENT1 <- as.numeric(rawData$PARENT1 == "Yes")

rawData$KIDSDRIV <- as.numeric(rawData$KIDSDRIV)
rawData$AGE <- as.numeric(rawData$AGE)
rawData$HOMEKIDS <- as.numeric(rawData$HOMEKIDS)
rawData$YOJ <- as.numeric(rawData$YOJ)
rawData$INCOME <- as.numeric(rawData$INCOME)
rawData$HOME_VAL <- as.integer(rawData$HOME_VAL)
rawData$BLUEBOOK<- as.integer(rawData$BLUEBOOK)
rawData$TIF <- as.integer(rawData$TIF)
rawData$OLDCLAIM <- as.integer(rawData$OLDCLAIM)
rawData$CLM_FREQ <- as.integer(rawData$CLM_FREQ)
rawData$MVR_PTS <- as.integer(rawData$MVR_PTS)
rawData$CAR_AGE <- as.integer(rawData$CAR_AGE)
```

```{r}
summary(select(rawData, -INDEX))
```

## Missing values

The rawData data frame contains 4506 missing values.
```{r}
nrow(rawData[is.na(rawData),])
```

The following variables in the dataset have missing values: **TARGET_FLAG**, **TARGET_AMT**, **AGE**, **YOJ**, and **CAR_AGE**. It is important to address these missing values before proceeding with any analysis.
```{r}
colSums(is.na(rawData))
```

Replace missing values in the YOJ column of the rawData data frame with **0**.
```{r}
rawData$YOJ <- ifelse(is.na(rawData$YOJ), 0, rawData$YOJ)
```
AGE
```{r}
head(rawData)
```

## Box-plot

```{r}
SEX_df <- rawData %>% 
  filter(!is.na(INCOME)) %>%
        group_by(SEX) %>% 
        summarise(INCOME = mean(INCOME))
SEX_df
```

```{r}
ggplot(rawData, aes(x = SEX, y = INCOME, color = SEX, fill = SEX)) +
  geom_bar(data = SEX_df, stat = "identity", alpha = .3) + 
    ggrepel::geom_text_repel(aes(label = HOME_VAL), color = "black", size = 2.5, segment.color = "grey") +
      geom_point() +
        guides(color = "none", fill = "none") +
        theme_bw() +
        labs(
          title = "Income by Sex",
          x = "Sex",
          y = "Income"
        )
```





